Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?

### 2021-06-24 docs/00-manuscripts/ieee/ieee-manuscript.tex
optimize data & study area

Airborne hyperspectral data with a spatial resolution of one meter and 126 spectral bands was available for four Monterey Pine (\textit{Pinus radiata D. Don}) plantations in northern Spain.
The trees in the plots suffer from infections of pathogens such as \textit{Diplodia sapinea (Fr.) Fuckel}, \textit{Fusarium circinatum Nirenberg \& O'Donnell}, \textit{Armillaria mellea (Vahl) P. Kumm}, \textit{Heterobasidion annosum (Fr.) Bref}, \textit{Lecanosticta acicola (Thüm) Syd.} and \textit{Dothisthroma septosporum (Dorogin) M. Morelet} causing a spread of cankers or defoliation \cite{mesanza2016, iturritxa2017}.
The last two fungi are mainly responsible for the foliation loss of the trees analyzed in this study \cite{iturritxa2014}.
In-situ measurements of defoliation of trees (serving as a proxy for tree health) were collected by visual inspection of experts in the field to serve as the response variable \textit{defoliation} which ranges from 0 - 100 (in \%) (\autoref{fig:defol-distr}).

@@

The \textit{Pinus radiata} plots of this study, namely Laukiz1, Laukiz2, Luiando and Oiartzun, are located in the northern part of the Basque Country (\autoref{fig:study_area}).
Oiartzun has the most observations (n = 559) while Laukiz2 shows the largest area size (1.44 ha).
All plots besides Luiando are located within 100 km from the coast (\autoref{fig:study_area}).
In total 1808 observations are available (Laukiz1 = 559, Laukiz2 = 451, Luiando = 301, Oiartzun = 497).
Field surveys were conducted in September 2016 by experienced forest pathologists.
Defoliation was measured via visual inspection using 5\% intervals with the help of a dedicated score card.
For Laukiz2, values at three height levels (bottom, mid, top) were available and averaged into an overall defoliation value, leading to values outside of the 5\% interval of the other three plots (e.g. 8.33 \%).
Estimating the human observer error when assessing defoliation is an issue which is being discussed since many years \cite{innes1993}.
Even though no estimation error was recorded in this study, \cite{maclean1982} estimated human observer errors when assessing defoliation to range between 7\% - 18\%.

@@

The airborne hyperspectral data was acquired during two flight campaigns which took place at noon on September 28th and October 5th 2016.
The images were taken by an AISA EAGLE-II sensor.
All preprocessing steps (geometric, radiometric, atmospheric) were conducted by the \ac{ICGC}.
The first four bands were corrupted, leaving 122 bands with valid information.
Additional metadata information is available in \autoref{tab:hyperparameter_metadata}.

>>>>

Airborne hyperspectral data with a spatial resolution of one meter and 126 spectral bands was available for four Monterey Pine (\textit{Pinus radiata D. Don}) plantations in northern Spain.
The trees in the plots suffer from infections of pathogens such as \textit{Diplodia sapinea (Fr.) Fuckel}, \textit{Fusarium circinatum Nirenberg \& O'Donnell}, \textit{Armillaria mellea (Vahl) P. Kumm}, \textit{Heterobasidion annosum (Fr.) Bref}, \textit{Lecanosticta acicola (Thüm) Syd.} and \textit{Dothisthroma septosporum (Dorogin) M. Morelet} causing (among others) needle blight, pitch canker and root diseases \cite{mesanza2016, iturritxa2017}.
The first two fungi are mainly responsible for the foliation loss of the trees analyzed in this study \cite{iturritxa2014}.
In-situ measurements of defoliation of trees (serving as a proxy for tree health) were collected by visual inspection of experts in the field to serve as the response variable \textit{defoliation} which ranges from 0 - 100 (in \%) (\autoref{fig:defol-distr}).

@@

The \textit{Pinus radiata} plots of this study, namely Laukiz1, Laukiz2, Luiando and Oiartzun, are located in the northern part of the Basque Country (\autoref{fig:study_area}).
Oiartzun consists of most tree observations (n = 559) while Laukiz2 shows the largest area size (1.44 ha).
All plots besides Luiando are located within 100 km from the coast (\autoref{fig:study_area}).
In total 1808 observations are available (Laukiz1 = 559, Laukiz2 = 451, Luiando = 301, Oiartzun = 497).
Field surveys were conducted in September 2016 by experienced forest pathologists.
Defoliation was measured via visual inspection using 5\% intervals with the help of a dedicated score card.
For Laukiz2, values at three height levels (bottom, mid, top) were available and averaged into an overall defoliation value, leading to values outside of the 5\% interval of the other three plots (e.g. 8.33 \%).
Estimating the human observer error of such surveys when assessing defoliation is an issue which is being discussed since many years \cite{innes1993}.
Even though no estimation error was recorded in this study, \cite{maclean1982} estimated human observer errors when assessing defoliation to range between 7\% - 18\%.

@@

The airborne hyperspectral data was acquired during two flight campaigns which took place at noon on September 28th and October 5th 2016.
Images were taken by an AISA EAGLE-II sensor.
All preprocessing steps (geometric, radiometric, atmospheric) were conducted by the \ac{ICGC}.
The first four bands were corrupted, leaving 122 bands with valid information.
Additional metadata information is available in \autoref{tab:hyperparameter_metadata}.
